---
title: "Reproducible research"
author: "Renukaneeli"
date: "10/17/2020"
output: html_document
---
## 1.Code for reading in the dataset and/or processing the data

```{r, echo=TRUE}
activity <- read.csv("activity.csv")
activity$date <- as.POSIXct(activity$date)
weekday <- weekdays(activity$date)
activity <- cbind(activity,weekday)

summary(activity)
colnames(activity)
```

## 2.Histogram of the total number of steps taken each day
```{r, echo=TRUE}
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
hist(activity_total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "green", ylim = c(0,20), breaks = seq(0,25000, by=2500))
```

## 3.Mean and median number of steps taken each day
```{r, echo=TRUE}
mean(activity_total_steps$steps)
median(activity_total_steps$steps)
```

## 4.Time series plot of the average number of steps taken
```{r, echo=TRUE}
average_daily_activity <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")
plot(average_daily_activity$interval, average_daily_activity$mean, type = "l", col="darkblue", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
```

## 5.The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo=TRUE}
average_daily_activity[which.max(average_daily_activity$mean), ]$interval
```

## 6.Code to describe and show a strategy for imputing missing data

```{r, echo=TRUE}
sum(is.na(activity$steps)) 
```

## 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r, echo=TRUE}

imputed_steps <- average_daily_activity$mean[match(activity$interval, average_daily_activity$interval)]

activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")

hist(total_steps_imputed$daily_steps, col = "darkblue", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
```

## 8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r, echo=TRUE}
activity$date <- as.Date(activity$date) 
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
activity_by_date <- aggregate(activity$steps~activity$interval + activity$datetype, activity, mean)
activity_by_date <- aggregate(steps~interval + datetype, activity, mean) 
```


```{r, echo=TRUE}
names(activity_by_date) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps ~ interval | day, activity_by_date, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")

```


